---
title: "lab 03"
author: "Aurielle Young, MPH"
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook provides completed code for lab 03.

## Dependencies
This notebook requires the following packages to complete the assignment:

```{r load-packages}
# tidyverse packages
library(ggplot2)      # statis mapping

# spatial packages
library(mapview)      # preview spatial data
library(sf)           # spatial tooks

# other packages
library(here)         # file path management 
library(RColorBrewer) # color brewer palettes
```

## Load Data
We will load a custom function:

```{r load-data}
source(here("source", "map_breaks.R"))
```

We are going to use `map_breaks` function.

A description of what your code should accomplish.
This notebook requires the following data from the `data/` folder.

```{r p1-q1}
# city boundary
stl_boundary <- st_read(here("data", "STL_BOUNDARY_City.geojson"))

# water layers
il_hydro <- st_read(here("data", "IL_HYDRO_Mississippi.geojson"))
stl_hydro <- st_read(here("data", "STL_HYDRO_AreaWater.geojson"))

# highways
highways <- st_read(here("data", "STL_TRANS_PrimaryRoads"))

# owner occupied housing
housing <- st_read(here("data", "STL_HOUSING_OwnerOccupied.geojson"))
```

## Part 1
The `mapview()` was used to explore the data. All layers are polygon data except for the highways which are line data.

## Part 2
The following code is used to create the map breaks using the `fisher` approach and then adds layers for the water features and highway to our census track data for owner occupied housing.

```{r housing-map} 
## create breaks
housing <- map_breaks(housing, var = "pct_owner_occupied", newvar = "map_breaks",
                      style = "fisher", classes = 5, dig_lab = 2)

## map binned data
p1 <- ggplot() +
  geom_sf(data = stl_boundary, fill = "#ffffff", color = NA) +
  geom_sf(data = housing, mapping = aes(fill = map_breaks)) +
  geom_sf(data = il_hydro, fill = "#d4f1f9") +
  geom_sf(data = stl_hydro, fill = "#d4f1f9") +
  geom_sf(data = highways, color = "#000000") +
  geom_sf(data = stl_boundary, fill = NA, color = "#2a2a2a", size = .6) +
  scale_fill_brewer(palette =  "Greens", name = "% Owner Occupied") +
  labs(
      title = "% of Housing Occupied by Homwowners",
      subtitle = "City of St. Louis 2010",
      caption = "Data via the U.S. Census Bureau\nMap by Aurielle Young, MPH"
  )
  
p1
```

Map shows that rates of ownership are highest in the southwest portion of St. Louis.

Will save the map using code:

```{r, save-plot}
ggsave(plot = p1, filename = here("results", "owner_occupied_housing.png"))
```

The map is now saved!
